<link rel="stylesheet" href="http://netdna.bootstrapcdn.com/bootstrap/3.0.3/css/bootstrap.min.css">
<style type="text/css"> body {padding: 10px 30px 10px 30px;} table,th, td {text-align: center;} </style>


Financial Data Segmentation: Factor Analysis
========================================================

**Arina Bobrova, Dennis Maroulis, Tavleen Kaur, Ville Herlin**
```{r echo=FALSE}
#Temp comment to fix Knit HTML issue:
#library(knitr)
#knit2html("/Users/ArinaB/Documents/Arina MBA/INSEAD-16D/P3/Big Data Analytics/AnalyticsProject/doc/FinancialDataFactorAnalysis.Rmd")
```

Business Decisions
---------------------------------------------------------
Analyze the correlation between different companies in the Vietnamise stock market.


The Data
--------------------------------------------------------------

We analyze 900 companies based on 17 different variables including market cap, EBITDA, PE, e.t.c. listed in the vietnamise stock market

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
ProjectDataFactor = data.matrix(ProjectDataFactor)
ProjectData = data.matrix(ProjectData)
```

This is how the first `r  min(max_data_report, nrow(ProjectData))` data looks:
<br>

<div class="row">
<div class="col-md-40">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectDataFactor,2))
show_data = show_data[1:min(max_data_report, nrow(ProjectData)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1920, height=min(400,27*(nrow(show_data)+1)),frozenColumns=1,allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>

Our Approach
---------------------------------------------------

#### Factor Analysis in 6 steps

1. Confirm the data in metric 

2. Decide whether to scale or standardize the data

3. Check the correlation matrix to see if Factor Analysis makes sense

4. Develop a scree plot and decide on the number of factors to be derived

5. Interpret the factors (consider factor rotations - technical but useful)

6. Save factor scores for subsequent analyses

Let's follow these steps.


The data we use here have the following descriptive statistics: 

<div class="row">
<div class="col-md-40">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
show_data = data.frame(round(my_summary(ProjectDataFactor),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1920, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>

#### Step 2: Decide whether to scale or standardize the data

We scaled the data because of the different magnitude of the numbers in the data under consideration.

```{r, results='asis'}
ProjectDatafactor_scaled=apply(ProjectDataFactor,2, function(r) {if (sd(r)!=0||is.na(r)==FALSE) res=(r-mean(r))/sd(r) else res=0*r; res})
```

Below is the summary statistics of the scaled dataset:

<br>


<div class="row">
<div class="col-md-40">
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}

show_data = data.frame(round(my_summary(ProjectDatafactor_scaled),2))
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1920, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>

<br>
As expected all variables have mean 0 and standard deviation 1. 



#### Step 3:  Check correlation matrix to see if Factor Analysis makes sense

This is the correlation matrix of the `r length(factor_attributes_used)` original variable we use for factor analysis. Some of the variables have strong correlations. The most obvious ones are between size meters such as revenue and value metrics such as market capitalization. Some interesting observations can be made from correlations alone. For instance dividend yield does not significantly affect valuation or depend on size. P/E does not depend on size. 

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
thecor = round(cor(ProjectDataFactor),2)
colnames(thecor)<-colnames(ProjectDataFactor)
rownames(thecor)<-colnames(ProjectDataFactor)
## printing the result in a clean-slate table
#cat(renderHeatmapX(thecor, border=1,center = 0,vrange_up = 1, vrange_down = 1))
#cat(renderHeatmapX(thecor, border=1))

show_data = data.frame(round(thecor,2))
#show_data=cbind(rownames(thecor),show_data)
m1<-gvisTable(show_data,options=list(width=1920, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE))
print(m1,'chart')
```


#### Step 4. Develop a scree plot and decide on the number of factors to be derived

Based on the cumulative explained variance we use six components that explain 67% of the total variance.

<br>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Component",1:ncol(UnRotated_Factors),sep=" ")
#```

#```{r echo=FALSE, comment=NA, warning=FALSE, #error=FALSE,message=FALSE,results='asis'}
#library("FactoMineR")
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table

row=1:nrow(Variance_Explained_Table)
name<-paste("Component No:",row,sep="")
Variance_Explained_Table<-cbind(name,Variance_Explained_Table)
Variance_Explained_Table<-as.data.frame(Variance_Explained_Table)
colnames(Variance_Explained_Table)<-c("Components", "Eigenvalue", "Percentage_of_explained_variance", "Cumulative_percentage_of_explained_variance")

m<-gvisTable(Variance_Explained_Table,options=list(width=1200, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'),formats=list(Eigenvalue="#.##",Percentage_of_explained_variance="#.##",Cumulative_percentage_of_explained_variance="#.##"))
print(m,'chart')
```
<br> <br>


```{r Fig1, echo=FALSE, comment=NA, results='asis', message=FALSE, fig.align='center', fig=TRUE}

eigenvalues  <- Variance_Explained_Table[,2]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
Line         <- gvisLineChart(as.data.frame(df), xvar="components", yvar=c("eigenvalues","abline"), options=list(title='Scree plot', legend="right", width=900, height=600, hAxis="{title:'Number of Components', titleTextStyle:{color:'black'}}", vAxes="[{title:'Eigenvalues'}]",  series="[{color:'green',pointSize:3, targetAxisIndex: 0}]"))
print(Line, 'chart')
```


```{r echo=FALSE, comment=NA, warning=FALSE,message=FALSE,results='asis'}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used

```

#### 5. Interpret the factors 

For our data, the selected factors look as follows after the rotation: 

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}

Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Component",1:ncol(Rotated_Factors),sep=" ")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

show_data <- Rotated_Factors 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
<br> <br>

Component 1 mainly consists of size metrics from recent performance and value metrics. Component 2 has some more longer time size metrics. It Component 3 consists of two debt capacity metrics. Component 4 combines historical and current p/e ratios. Component 5 has two variables that measure the valuation in relation to performance. Component 6 groups two metrics of debt to assets.

Components 1 and 2 have somewhat similar loadings because all the variables correlate strongly with each other. It might not make sense to use both of them.

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)

show_data <- Rotated_Factors_thres 
#show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
<br> <br>


#### Step 6. Save factor scores for subsequent analyses

For our data, using the rotated factors we selected, we can create a new dataset where our observations are as follows (for the first 50 companies):

<div class="row">
<div class="col-md-40">

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
NEW_ProjectData <- round(Rotated_Results$scores[,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste("Derived Variable (Factor)",1:ncol(NEW_ProjectData),sep=" ")

if (factors_selected >=2){ 
  
  show_data <- as.data.frame(NEW_ProjectData) 
  show_data = show_data[1:min(max_data_report,nrow(show_data)),]
  row<-rownames(show_data)
  dfnew<-cbind(row,show_data)
  change<-colnames(dfnew)
  change[1]<-"Observation"
  colnames (dfnew)<-change
  m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),frozenColumns=1,allowHTML=TRUE,page='disable'))
  print(m1,'chart')
  
  } else {
    print(xtable(NEW_ProjectData, caption="Only 1 derived variable generated", digits=3), type="html",html.table.attributes="class='table table-striped table-hover table-bordered'", caption.placement="top", comment=FALSE, include.rownames=TRUE, include.colnames=TRUE)
    }
```
</div>
</div>
<br> <br>

Further Ideas
---------------------------------------------------------------

This framework can be extended to work as a tool for screening companies. It would make sense to add a lot more variables. Then components can be compared with other markets to test hypothesis along the lines “does debt capacity matter more or less for valuation in these markets?” The framework can also be extended for back testing investment strategies. 
