<link rel="stylesheet" href="http://netdna.bootstrapcdn.com/bootstrap/3.0.3/css/bootstrap.min.css">
<style type="text/css"> body {padding: 10px 30px 10px 30px;} table,th, td {text-align: center;} </style>


Boats Segmentation: Factor Analysis
========================================================

**Team names**


Business Decisions
---------------------------------------------------------
your text...


The Data
--------------------------------------------------------------


your text...

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
ProjectDataFactor = data.matrix(ProjectDataFactor)
#ProjectData = data.matrix(ProjectData)
```

This is how the first `r min(max_data_report, nrow(ProjectData))` data looks:
<br>

<div class="row">
<div class="col-md-6">
```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
show_data = data.frame(round(ProjectDataFactor,2))
show_data = show_data[1:min(max_data_report,nrow(show_data)),]
row<-rownames(show_data)
dfnew<-cbind(row,show_data)
change<-colnames(dfnew)
change[1]<-"Variables"
colnames (dfnew)<-change
m1<-gvisTable(dfnew,options=list(showRowNumber=TRUE,width=1220, height=min(400,27*(nrow(show_data)+1)),allowHTML=TRUE,page='disable'))
print(m1,'chart')
```
</div>
</div>
<br> <br>


Our Approach
---------------------------------------------------

#### Factor Analysis in 6 steps

1. Confirm the data in metric 

2. Decide whether to scale or standardize the data

3. Check the correlation matrix to see if Factor Analysis makes sense

4. Develop a scree plot and decide on the number of factors to be derived

5. Interpret the factors (consider factor rotations - technical but useful)

6. Save factor scores for subsequent analyses

Let's follow these steps.


The data we use here have the following descriptive statistics: 

